Lightness constancy under fixed geometry

### 2017-10-15 Draft1/draft1.tex
more edits

\title{Lightness constancy under fixed geometry}
\abstract{Human visual system has the ability to stably perceive the color of an object under variations in the reflected light due to surround factors. This is a behaviorally significant invariance detection ability whose biological mechanism remains unknown. To understand the computational principles that can lead to such invariant perception, here we study the perception of object lightness under spectral variations. Lightness is the attribute of the 3D color space that determines the brightness of a color. Specifically, we study the effects of variations in reflectance and illumination spectra of the objects and lights sources in a scene on the perceived lightness of an object. We have developed a software to generate naturalistic hyper-spectral images of 3D scenes with precise control over the geometrical and spectral properties of the objects and lights in the scene. The software also produces the response of the early visual system to these hyper-spectral images. We label the images with the lightness of a specific object in the scene and use supervised learning methods to identify the computations that lead to stable lightness perception. Our analysis shows that for simple spectral variations, like the ones that allow variations only in the spectrum of the target object, or of the background object, or of the light source, the standard lightness can be recovered by through transformations of the reflected light. When all spectra are allowed to vary simultaneously, while it is not easy to recover the lightness through simple transformations, a decoding sheme that compares the light from the target and the surround, and properly adds the response of the L,M and S cones can recover the lightness within about 15\% RMSE.}

@@

\section{Introduction}
The color of an object has important behavioral implications. This important property of an object is perceived from the light reflected off the object surface. The reflected light depends not just on the properties intrinsic to the object, such as the object's surface reflectance, its texture, material properties and geometry (shape, size, orientation), but also on extrinsic factors that depend on the object's surround (Fig. \ref{fig:introSchematic}). These extrinsic factors are: the surface reflectance, texture and material properties of the objects in the surround, the illumination spectrum of the sources of light illuminating the object, the geometry, relative position and pose of the objects and light sources in the surround, and the position of the observer. As these extrinsic factors vary considerably from scene to scene, the challenge faced by the visual system is to recover a stable perceptual representation of the object's surface reflectance, invariant of the extrinsic factors. The ability of the visual system to extract such a stable representation, its perceived color, is called color constancy. We are interested in identifying the computations that lead to such constancy.

In this work, we identify the computations that lead to stable lightness perception. Lightness is the attribute that determines the brightness of a color. A color with a low lightness is nearly black, while one with a high lightness is the pure, saturated color. For achromatic objects, lightness is the property that changes from black to gray to white. We study the effects of variations in the reflected light on the perceived lightness.

% Figure 1: Introduction
\begin{figure}[h]
\centering
\begin{subfigure}{0.4 \textwidth}
		\centering
        \includegraphics[width=\textwidth]{../Figures/Figure1/Figure1_a.png}
        \caption{}
        \label{fig:introSchematic}
    \end{subfigure}
    \begin{subfigure}{0.45 \textwidth}   
        \includegraphics[width=\textwidth]{../Figures/Figure1/Figure1_b.jpeg}
        \caption{}
        \label{fig:introExampleFigure}
    \end{subfigure}
    \label{introFigure}
    \caption{(a) {\bf Color Constancy:} The light reflected off an object and incident on our eyes, depends both on the surface reflectance of the object $\left( R(\lambda,x) \right)$ and the illumination spectrum of the light source $\left(E(\lambda,x)\right)$. Additionally, the reflected light also depends on context the object lies in. Thus a variation in the reflected light can result from either a change in the object surface reflectance, or a change in the illumination spectrum of the light source, or change in the context, or all of these. Color constancy refers to the ability of the human visual system to perceive the color of the object stably under variations coming from these factors. (b) {\bf Lightness Constancy Under Spectral Variations:} In this work we study lightness constancy under variations in the spectra of the light sources and the objects. In Fig.\ref{fig:introExampleFigure}, we show 5 images generated at three different lightness levels of the target object (spherical object in the center). Along the rows we have fixed the lightness level of the target object, while we vary the spectra of the target object, the background objects and the light sources. (For illustration, in fig.\ref{fig:introExampleFigure}, We have fixed the shape of the target object reflectance spectra along the column.) Our aim is to identify computations that can recover the target object lightness from such stimuli.} 
\end{figure}

The variations in the reflected light can result from either geometrical factors (like shape, size, relative position of objects, etc.) or spectral factors (such as the illumination spectra, surface reflectance, diffusivity, etc.). Here, we focus on the spectral factors. Specifically, we study the effects of variations in the surface reflectance spectra and the illumination spectra on lightness perception  (Fig.\ref{fig:introExampleFigure}). We are interested in identifying the computations that can lead to stable lightness perception.

\begin{comment}
\begin{figure}
\centering
\includegraphics[width=\textwidth]{approach.png}
\caption{{\bf Approach:} We have employed a supervised learning approach to study color constancy. We first generate a large dataset of hyperspectral images labeled with the lightness level of the target object. We generate images at multiple lightness levels, with images at the same lightness level having variable object reflectance spectra and illumination spectra. These hyperspectral images are used to simulate the response of the visual system. We use isetbio toolbox to model the optics and physiology of early vision and produce the response of the cones in the retina. We use supervised computational learning methods on this labeled dataset of cone-responses to estimate the lightness of the target object. The performance of the supervised method is calculated using a separate test dataset.} 
\label{fig:approach}
\end{figure}
\end{comment}

We employ a supervised learning approach to understand this phenomenon. We generate a database of hyperspectral images of naturalistic scenes labeled by the lightness of a specific target object in the scene. The images are labeled with perceived lightness of a uniform surface having the target object reflectance placed under the standard daylight spectrum $D_{65}$. We generate hundreds of hyperspectral images at several fixed values of the target object standard lightness. At a particular lightness level, we allow variations in the surface reflectance spectra of the objects (both the target and the surround objects) and the illumination spectra of the light sources (Fig. \ref{fig:introExampleFigure}.

In this work, we study three cases resulting from the combination of such spectral variations (Fig.\ref{fig:studiedCases}). These cases are: 1. Variations in target object spectra (Fig.\ref{fig:targetVarying}) with fixed illumination and background surface reflectance spectra, 2. Variations in target object spectra and the light source spectra, with fixed background surface reflectance spectra (Fig.\ref{fig:targetIlluminantVarying}), and 3. Variations in all spectra (target object, background object and light source) (Fig.\ref{fig:allSpectraVarying}). The effects of other types of spectral variations can be explained in terms of one of these cases (see supplementary information).

% Figure 2: Cases Studied
\begin{figure}
\centering
	\begin{subfigure}[b]{0.33 \textwidth}
		\includegraphics[width=\textwidth]{../Figures/Figure2/Figure2_a.jpeg}
		\caption{Case 1}
 		\label{fig:targetVarying}
	\end{subfigure}
	\begin{subfigure}[b]{0.33 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure2/Figure2_b.jpeg}
        \caption{Case 2}
        \label{fig:targetIlluminantVarying}
    \end{subfigure}
	\begin{subfigure}[b]{0.33 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure2/Figure2_c.jpeg}
        \caption{Case 3}
        \label{fig:allSpectraVarying}
    \end{subfigure}    
    \caption{{\bf Three types of spectral variations studied in this work:} sRGB rendition of images studied in our work. The numbers on the left are the standard lightness level of the target object. We show 5 images for each case at 3 lightness levels. The analysis was performed with images at 10 lightness levels. We studied three types of spectral variations. (a) Case 1: Target reflectance spectra variable, Background reflectance spectra fixed, illuminant spectra fixed (b) Case 2: Target reflectance spectra variable, Background reflectance spectra fixed, illuminant spectra variable (c) Case 3: Target reflectance spectra variable, Background reflectance spectra variable, illuminant spectra variable.}
\label{fig:studiedCases}
\end{figure}

After generating databases of labeled hyperspectral images for each one of these cases, we simulate the response of the cones in the retina cone using an accurate model of the eye. These cone-responses datasets, that are labeled with the standard lightness of the target object, are then analyzed using supervised computational learning methods.

We have employed three supervised computational learning methods for estimating the target object lightness: linear regression, a generalized SVD regression and accuracy maximization analysis (AMA). The linear regression has been performed on the cones corresponding to the center pixel of the target object. The other two methods take the response of all the cones in the model retina. While, SVD regression a linear transformation to estimate the lightness, AMA uses a non-linear transform of the data. These methods identify the computations that need to be performed on the cone-responses to estimate the target object lightness. Finally, we evaluate the performance of the computational methods on a test set of images.

@@

We show that for cases that allow variation either only in the target surface reflectance or the background surface reflectance or the spectral power distribution of the source of light, the standard lightness ($Y_{65}$) can be recovered by simple manipulations of the reflected light captured by the visual system. For case 1, (variations only in target obejct reflectance spectra), $Y_{65}$ can be recovered directly from the light reflected off the target object. For case 2, where both the target object reflectance and light source spectra vary, $Y_{65}$ can be recovered by from the contrast between the target and the background. When all the spectral features in the scene are allowed to change (case 3), the standard lightness can not be recovered from a simple manipulations of the light captured by the visual system. Here, we have used two supervised learning methods with Bayesian estimators to estimate standard lightness of the target object. Using these methods, we can recover the lightness of the target object with $\sim 15\%$ relative root mean square error.

Our analysis shows that the optimal receptive fields for estimating the lightness of the target object show a center surround structure, supporting a comparison between the light coming from the target object and its surround. Additionally, lightness estimation invovles a weighted sum of the L, M and S cone responsese where the receptive fields give more weight to the L and M cone responses compared to S cone responses. Comparing the performace of the receptive fields of one case on the cone response of other cases shows that the receptive field of the complex cases perform well on images of simpler cases.  We discuss the implications of such receptive fields on lightness constancy.

\section{Methods}
\subsection{Image database} \label{method:VirtualWorld}
In this section we briefly describe our image generation software that we call Virtual world color constancy (VWCC). The details of the software are provided in the supplementary information section. To generate a virtual scene, we first select a 3D scene, a target object and a light source from the set of predefined scenes and objects of the VWCC software. The target object and the light source are inserted at predefined positions in the 3D scene. Fig.\ref{fig:3DScene} shows a 2D sRGB rendition of the 3D scene with the inserted target object (green sphere in the middle) used in our analysis. (The light source in not visible.) Next, we assign a surface reflectance spectrum and an illumination spectrum to each object and light source in the scene. For the target object the surface reflectance is assigned such that the perceived lightness of the target object under D65 daylight spectrum has a fixed value. Once these spectral properties have been assigned, we render a 320x240 pixel$^2$ 2d hyperspectral images of the scene at 31 frequencies (equal linear spacing between 400nm - 700nm) using a graphics rendering system called Mitsuba \cite{jakob2015mitsuba}. We generate hundreds of such scenes and the corresponding hyperspectral images at 10 linearly spaced lightness levels between 0.2 and 0.6. For the images at a fixed lightness level, the reflectance and illumination spectra are assigned according to the spectral variations allowed under each case. Next, we crop a 51x51 pixel$^2$ part of the hyperspectral image containing the target object for further analysis (\ref{fig:croppedImage}). Fig. \ref{fig:studiedCases} shows examples of such cropped image at 3 lightness levels for the three cases considered in this work.

% Figure 3: Methods
\begin{figure}
\centering
\begin{subfigure}[b]{0.25 \textwidth}
		\centering
        \includegraphics[width=\textwidth]{../Figures/Figure3/Figure3_a.png}
        \caption{A typical 3D Scene}
        \label{fig:3DScene}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.19 \textwidth}   
    \hspace{0.1 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure3/Figure3_b.png}
        \caption{Cropped Image}
        \label{fig:croppedImage}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
    %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.19 \textwidth}
    \hspace{0.1 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure3/Figure3_c.png}
        \caption{Optical Image}
        \label{fig:croppedImageWithMosaic}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.2 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure3/Figure3_d.png}
        \caption{LMS cone contrast}
        \label{fig:coneContrast}
    \end{subfigure}

    \label{fig:sceneWithCroppedImage}
    \caption{{\bf Generating labeled dataset for computational analysis:}  We employ a supervised learning approach in our study. We generate a database of visual system response to images labeled with the lightness of the target object. This is done as follows: (a) We define a 3D virtual scene where we can assign the geometrical and spectral properties of the objects. In this 3D scene, we place a target object (here the spherical object) in the camera field of view and assign a fixed reflectance spectrum to the target object. Then we render a hyperspectral image of the scene. (b) A part of the image, containing the target object, is cropped from the image. (c) Using an accurate model of the front end of the visual system, we generate the response of the retinal cones to the cropped image. Fig.\ref{fig:croppedImageWithMosaic} shows the optical image incident on the cone mosaic and the location and the identity of the cones (L cones: Red, M cones: Green, S cones: Blue). See text for the details of the model. (d) For each cone type, we interpolate the response to get the cone-response of all the three types at each location. These responses are contrast normalized (see text) to the get the input analysis.}
\end{figure}

\subsection{Model of Visual System} \label{method:Isetbio}
We simulate the visual system response using \href{http://isetbio.org}{Isetbio} software toolbox. Isetbio simulates a model of the visual system accounting for cornea, pupil size, lens and physiological optics. We use the hyperspectral cropped image (fig.\ref{fig:croppedImage}) as the optical stimulus to simulate the response of the cones in the retina. The cropped image patch is assumed to cover a $1^{\circ}$ field of view at a distance of 1 m. The retinal cone mosaic is modeled as a square grid with 51x51 cones in the l:m:s ratio 0.6:0.3:0.1. The response of the cones are measured in terms of the number of isomerizations in the cone segment ({\it over what time ??}). Since, S cones response are much lower than of L and M cones, to model the amplification of the S cone response downstream, we rescale the cone response functions by the total area under the response curves. This leads to comparable number of isomerizations in the L, M and S cones. Further, to get the response of each cone class at each poistion on the retinal mosaic, we interpolate the response of the cones for each cone class. Thus we get a $51 \times 51$ cone response image for each cone class.

@@

To overcome this  challenge, we have developed an image rendering software pipeline that allows us to specify a virtual, but naturalistic, statistical model of visual scenes with parametric control over factors in natural scenes. The software uses a physically-based image rendering system that accurately accounts for the interactions of light with matter in 3D scenes. It then renders a hyperspectral 2D image of each modeled visual scene. Wherever possible, we have incorporated statistical models of the variation of specific scene parameters based on observed data in natural scenes. In addition, our software simulates the responses of retinal cones to the rendered hyperspectral images using a realistic model of the early human visual system. Below we provide the details of the software and the choices made to generate ensembles of scenes with corresponding images. Our software, which is written in MATLAB, is freely available through a public respository: \href{https://github.com/BrainardLab/VirtualWorldColorConstancy}{Virtual World Color Constancy}. It builds on several other open source projects, including \href{http://rendertoolbox.org}{RenderToolbox4} \cite{heasly2014rendertoolbox3}, \href{http://isetbio.org}{Isetbio} and \href{https://www.mitsuba-renderer.org}{Mitsuba} \cite{jakob2015mitsuba}. Below we refer to the software pipeline as VWCC (for Virtual World Color Constancy).

@@

After specifying the geometry and spectra, one of the inserted objects is chosen as the target object in the scene. A camera is inserted at a user specified position and then pointed at the target object. We ensure that the target object is not completely occluded by another object in the line of sight of the camera. Next, a predefined surface reflectance spectrum is assigned to the target object. This target surface reflectance sets the label for the virtual scene image. The scene is now rendered using the Mitsuba rendering package. The renderer produces 2D hyper-spectral images of the scene at specified wavelengths. Finally, these hyperspectral images are used to simulate retinal cone responses using the \href{https://github.com/isetbio}{ISETBIO} software.

@@

For the images used in this work, we first selected the ``{\it Library}'' base scene from the VWCC base scene set. We inserted the spherical ``{\it Big Ball}'' object in the base scene and called it the target object. We also inserted a light source with the same spherical ``Big Ball'' shape. Also, a camera is inserted the scene and pointed at the target object. The position, and size of the inserted object, light source and the camera was held fixed for all the images analyzed in this work, i.e., no geometrical manipulations were performed between the images. Next, we assigned illumination and surface reflectance spectra to the lights and objects in the scene. The spectral manipulations for the various cases studied in this work are described below. hyperspectral 2D images were rendered at 31 wavelengths from $400$nm to $700$nm spaced $10$nm apart at an image size $320\times 240$ pixel$^2$. Next, a $51 \times 51$ pixel$^2$ patch of the hyperspectral image was cropped around the target object and cone responses were calculated to these cropped images as described below.

>>>>

\title{Lightness constancy under fixed geometry}
\abstract{The light that reflects off an object surface and is captured by our eyes varies significantly with its context. But, the human visual system has the ability stably perceive the object's color invariant of the context. The mechanism of this behaviorally significant invariance detection ability remains largely unknown. To understand the computational principles that lead to such invariant perception, we study the perception of object lightness under spectral variations in 3D scenes. Specifically, we study the effects of variations in reflectance and illumination spectra in a scene on the perceived lightness of an object. We have developed a software that generates naturalistic hyper-spectral images of 3D scenes with precise control over the geometrical and spectral properties of the constituents that make a scene. We label these images with the lightness of a specific object in the scene. Next, we simulate the response of the retinal cones to these hyper-spectral images using an accurate model of the early visual system.  Using supervised learning methods on these labeled cone response we identify the computations that lead to accurate lightness estimation. We show that if only target, or background, or illumination spectra are allowed to vary, the standard lightness can be estimated through simple transformations of the cone responses. When all the spectra vary simultaneously, while it is not easy to recover the lightness through simple transformations, a decoding scheme that compares the light from the target and the surround, and properly adds the response of the L,M and S cones can recover the lightness within about 15\% RMSE.}

@@

\section{Introduction}
The color of an object has important behavioral implications. This important property of an object is perceived from the light reflected off the object surface. The reflected light depends not just on the properties intrinsic to the object, such as the object's surface reflectance, its texture, material properties and geometry (shape, size, orientation), but also on extrinsic factors that depend on the object's surround (Fig. \ref{fig:introSchematic}). These extrinsic factors, such as the surface reflectance, texture and material properties of the objects in the surround, the illumination spectrum of the sources of light illuminating the object, the geometry, relative position and pose of the objects and light sources in the surround, and the position of the observer, vary considerably from scene to scene. The challenge faced by the visual system is to recover a stable perceptual representation of the object's surface reflectance invariant of the extrinsic factors. The ability of the visual system to extract such a stable representation, its perceived color, is called color constancy. We are interested in identifying the computations that lead to such constancy.

In this work, we study lightness estimation of objects in naturalistic scenes using supervised learning methods. Lightness is the attribute that determines the brightness of a color. A color with a low lightness is nearly black, while one with a high lightness is the pure, saturated color. For achromatic objects, lightness is the property that changes from black to gray to white. Here, we study the effects of variations in the reflected light on the perceived lightness. These variations can result either due to geometrical factors (like shape, size, relative position of objects, etc.) or due to spectral factors (such as the illumination spectra, surface reflectance, diffusivity, etc.). We focus on spectral factors, specifically the effects of variations in the surface reflectance spectra and the illumination spectra (Fig.\ref{fig:introExampleFigure}). Our aim is to identify the computations that lead to accurate lightness estimation.

% Figure 1: Introduction
\begin{figure}
\centering
\begin{subfigure}{0.4 \textwidth}
		\centering
        \includegraphics[width=\textwidth]{../Figures/Figure1/Figure1_a.png}
        \caption{}
        \label{fig:introSchematic}
    \end{subfigure}
    \begin{subfigure}{0.45 \textwidth}   
        \includegraphics[width=\textwidth]{../Figures/Figure1/Figure1_b.jpeg}
        \caption{}
        \label{fig:introExampleFigure}
    \end{subfigure}
    \label{introFigure}
    \caption{(a) {\bf Color Constancy:} The light coming from an object to our eyes depends both on the surface reflectance of the object, $R(\lambda,x)$, and the illumination spectrum of the light source, $E(\lambda,x)$. Additionally, it depends on the context in which the object lies. The human visual system has the ability to discard any variations in the captured light due to extrinsic factors, and stably perceive the color of the object. Such invariant color perception is called color constancy. (b) {\bf Lightness Constancy Under Spectral Variations:} In this work we study lightness estimation. Fig.\ref{fig:introExampleFigure} shows images of an object (spherical object in the center) at three lightness levels. We show 5 images at each lightness level which vary in the spectra of the target object (sphere), the background objects and the light sources. (For illustration, in fig.\ref{fig:introExampleFigure}, we have fixed the shape of the target object reflectance spectra along the column.) Our aim is to identify computations that can accurately estimate the object lightness under such variations.} 
\end{figure}

\begin{comment}
\begin{figure}
\centering
\includegraphics[width=\textwidth]{approach.png}
\caption{{\bf Approach:} We have employed a supervised learning approach to study color constancy. We first generate a large dataset of hyper-spectral images labeled with the lightness level of the target object. We generate images at multiple lightness levels, with images at the same lightness level having variable object reflectance spectra and illumination spectra. These hyper-spectral images are used to simulate the response of the visual system. We use isetbio toolbox to model the optics and physiology of early vision and produce the response of the cones in the retina. We use supervised computational learning methods on this labeled dataset of cone-responses to estimate the lightness of the target object. The performance of the supervised method is calculated using a separate test dataset.} 
\label{fig:approach}
\end{figure}
\end{comment}

We employ a supervised learning approach to understand this phenomenon. We generate a database of hyper-spectral images of naturalistic scenes, labeled by the lightness of a specific target object in the scene. These images are labeled with the perceived lightness of the target object under standard daylight spectrum $D_{65}$ (see Methods). We generate hundreds of images with variations in the surface reflectance spectra of the objects and the illumination spectra of the light sources (Fig.~\ref{fig:introExampleFigure}). We study three cases resulting from the combination of such spectral variations (Fig.~\ref{fig:studiedCases}). These cases are: 1. Variation in target object spectrum (Fig.\ref{fig:targetVarying}) with fixed illumination and background, 2. Variations in both the target object spectrum and the light source spectra, with fixed background spectra (Fig.\ref{fig:targetIlluminantVarying}), and 3. Variations in all three spectra (target, background and light source) (Fig.\ref{fig:allSpectraVarying}). The effects of other types of spectral variations can be explained in terms of one of these cases (see Discussion/Supplementary information). Next, we simulate the response of the cones in the retina to these hyper-spectral images using an accurate model of the early visual system. Such labeled cone-responses datasets are then analyzed using supervised computational learning methods. We use three methods: linear regression, generalized SVD regression and accuracy maximization analysis (AMA) in our analysis. These methods identify the computations that need to be performed on the cone-responses to estimate the target object lightness. These methods identify the receptive fields (RFs) that optimal for lightness estimation. We evaluate the performance of the learnt RFs on a test set of images.

% Figure 2: Cases Studied
\begin{figure}
\centering
	\begin{subfigure}[b]{0.33 \textwidth}
		\includegraphics[width=\textwidth]{../Figures/Figure2/Figure2_a.jpeg}
		\caption{Case 1}
 		\label{fig:targetVarying}
	\end{subfigure}
	\begin{subfigure}[b]{0.33 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure2/Figure2_b.jpeg}
        \caption{Case 2}
        \label{fig:targetIlluminantVarying}
    \end{subfigure}
	\begin{subfigure}[b]{0.33 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure2/Figure2_c.jpeg}
        \caption{Case 3}
        \label{fig:allSpectraVarying}
    \end{subfigure}    
    \caption{{\bf Three types of spectral variations studied in this work:} sRGB rendition of hyper-spectral images similar to ones studied in our work. The numbers on the left indicate the standard lightness level of the target object. 5 images are shown at each lightness level. We studied three types of spectral variations. (a) Case 1: Target reflectance spectrum variable, illumination spectra fixed, background spectra fixed, (b) Case 2: Target spectrum variable, Illumination spectra variable, Background spectra fixed, (c) Case 3: Target spectrum variable, Illumination spectra variable, Background spectra variable.}
\label{fig:studiedCases}
\end{figure}

@@

We show that for cases that have spectral variation only in target reflectance or background reflectance or the illuminant, the standard lightness ($Y_{65}$) can be estimated by simple transformations of the cone responses. For variations only in target object reflectance spectra (case 1), object lightness can be recovered directly from the light reflected off the target object. For case 2, where both the target object reflectance and light source spectra vary, object lightness can be recovered from the contrast between the target and the background. When all the spectral features in the scene are allowed to change (case 3), the standard lightness can not be recovered simply from the cone response or the contrast. Here, we have used two supervised learning methods with Bayesian estimators to estimate standard lightness of the target object. Using these methods, we can recover the lightness of the target object with $\sim 15\%$ relative root mean square error.

Our analysis shows that the optimal receptive fields for estimating the lightness of the target object show a center surround structure, supporting a comparison between the light coming from the target object and its surround. Additionally, lightness estimation involves a weighted sum of the L, M and S cone responses where the receptive fields give more weight to the L and M cone responses compared to S cone responses. Comparing the performance of the RFs learnt using  one case on the cone response of other cases shows that the receptive field of the complex cases perform well on images of simpler cases. We discuss the implications of such receptive fields on lightness constancy.

\section{Methods}
\subsection{Image database} \label{method:VirtualWorld}
In this section we briefly describe our image generation software that we call Virtual world color constancy (VWCC). The details of the software are provided in the supplementary information section. We first select a 3D scene, a target object and a light source from the set of predefined scenes and objects of the VWCC software. The target object and the light source are inserted at predefined positions in the 3D scene. Fig.\ref{fig:3DScene} shows a 2D sRGB rendition of the 3D scene used in our analysis. The green sphere in the middle is the target object. The light source is not in the camera field of view. Next, we assign a surface reflectance spectrum and an illumination spectrum to each object and light source in the scene. For the target object the surface reflectance is assigned such that the perceived lightness of the target object under D65 daylight spectrum has a fixed value. Now we render a $320\times 240$ pixel$^2$ 2d hyper-spectral images of the scene at 31 frequencies (equal linear spacing between 400 nm - 700 nm) using a graphics rendering system called Mitsuba \cite{jakob2015mitsuba}. For each of the three case, We generate hundreds of such scenes and corresponding hyper-spectral images at 10 linearly spaced lightness levels between 0.2 and 0.6 (see below). We crop a 51x51 pixel$^2$ part of the hyper-spectral image containing the target object for further analysis (\ref{fig:croppedImage}). Fig. \ref{fig:studiedCases} shows examples of such cropped image at 3 lightness levels for the three cases considered in this work.

% Figure 3: Methods
\begin{figure}
\centering
\begin{subfigure}[b]{0.25 \textwidth}
		\centering
        \includegraphics[width=\textwidth]{../Figures/Figure3/Figure3_a.png}
        \caption{A typical 3D Scene}
        \label{fig:3DScene}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.19 \textwidth}   
    \hspace{0.1 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure3/Figure3_b.png}
        \caption{Cropped Image}
        \label{fig:croppedImage}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
    %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.19 \textwidth}
    \hspace{0.1 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure3/Figure3_c.png}
        \caption{Optical Image}
        \label{fig:croppedImageWithMosaic}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.2 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure3/Figure3_d.png}
        \caption{LMS cone contrast}
        \label{fig:coneContrast}
    \end{subfigure}
    \label{fig:sceneWithCroppedImage}
    \caption{{\bf Generating labeled dataset for computational analysis:}  For the supervised approach employed in this work, the labeled database is generated as follows: (a) A 3D virtual scene, a target object and a light source is selected from the VWCC library. The target object is placed in the camera field of view and assigned a reflectance spectrum with a specific lightness. After assigning spectra to other objects and lights sources in the scene, a hyper-spectral image of the scene is rendered using Mitsuba graphics rendering software. (b) A part of the image, containing the target object, is cropped from the image. (c) The response of the retinal cones to the cropped image is simulated using a model of the visual system. Fig.\ref{fig:croppedImageWithMosaic} shows the optical image incident on the cone mosaic and the location and the identity of the cones (L cones: Red, M cones: Green, S cones: Blue). See text for the details of the model. (d) The cone response are interpolated to get the response of all the three types at each location (demosaicing). Finally, the demosaic response is contrast normalized (see text).}
\end{figure}

\subsection{Model of Visual System} \label{method:Isetbio}
We simulate the visual system response using \href{http://isetbio.org}{Isetbio} software toolbox. Isetbio simulates a model of the visual system accounting for cornea, pupil size, lens and physiological optics. We use the hyper-spectral cropped image (fig.\ref{fig:croppedImage}) as the optical stimulus to simulate the response of the cones in the retina. The cropped image patch is assumed to cover a $1^{\circ}$ field of view at a distance of 1 m. The retinal cone mosaic is modeled as a square grid with 51x51 cones in the l:m:s ratio 0.6:0.3:0.1. The response of the cones are measured in terms of the number of isomerizations in the cone segment ({\it over what time ??}). Since, S cones response are much lower than of L and M cones, to model the amplification of the S cone response downstream, we rescale the cone response functions by the total area under the response curves. This leads to comparable number of isomerizations in the L, M and S cones. Further, to get the response of each cone class at each poistion on the retinal mosaic, we interpolate the response of the cones for each cone class. Thus we get a $51 \times 51$ cone response image for each cone class.

@@

To overcome this  challenge, we have developed an image rendering software pipeline that allows us to specify a virtual, but naturalistic, statistical model of visual scenes with parametric control over factors in natural scenes. The software uses a physically-based image rendering system that accurately accounts for the interactions of light with matter in 3D scenes. It then renders a hyper-spectral 2D image of each modeled visual scene. Wherever possible, we have incorporated statistical models of the variation of specific scene parameters based on observed data in natural scenes. In addition, our software simulates the responses of retinal cones to the rendered hyper-spectral images using a realistic model of the early human visual system. Below we provide the details of the software and the choices made to generate ensembles of scenes with corresponding images. Our software, which is written in MATLAB, is freely available through a public respository: \href{https://github.com/BrainardLab/VirtualWorldColorConstancy}{Virtual World Color Constancy}. It builds on several other open source projects, including \href{http://rendertoolbox.org}{RenderToolbox4} \cite{heasly2014rendertoolbox3}, \href{http://isetbio.org}{Isetbio} and \href{https://www.mitsuba-renderer.org}{Mitsuba} \cite{jakob2015mitsuba}. Below we refer to the software pipeline as VWCC (for Virtual World Color Constancy).

@@

After specifying the geometry and spectra, one of the inserted objects is chosen as the target object in the scene. A camera is inserted at a user specified position and then pointed at the target object. We ensure that the target object is not completely occluded by another object in the line of sight of the camera. Next, a predefined surface reflectance spectrum is assigned to the target object. This target surface reflectance sets the label for the virtual scene image. The scene is now rendered using the Mitsuba rendering package. The renderer produces 2D hyper-spectral images of the scene at specified wavelengths. Finally, these hyper-spectral images are used to simulate retinal cone responses using the \href{https://github.com/isetbio}{ISETBIO} software.

@@

For the images used in this work, we first selected the ``{\it Library}'' base scene from the VWCC base scene set. We inserted the spherical ``{\it Big Ball}'' object in the base scene and called it the target object. We also inserted a light source with the same spherical ``Big Ball'' shape. Also, a camera is inserted the scene and pointed at the target object. The position, and size of the inserted object, light source and the camera was held fixed for all the images analyzed in this work, i.e., no geometrical manipulations were performed between the images. Next, we assigned illumination and surface reflectance spectra to the lights and objects in the scene. The spectral manipulations for the various cases studied in this work are described below. hyper-spectral 2D images were rendered at 31 wavelengths from $400$nm to $700$nm spaced $10$nm apart at an image size $320\times 240$ pixel$^2$. Next, a $51 \times 51$ pixel$^2$ patch of the hyper-spectral image was cropped around the target object and cone responses were calculated to these cropped images as described below.
